Idea Development

### 2024-11-07 posts/Abortion/abortion.qmd
ideas

---
title: "Abortion"
date: 11/03/2024
date-modified: last-modified
date-format: full
categories: [Others]
website:
  comments: 
    utterances:
      repo: quarto-dev/quarto-web
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
bibliography: references.bib
---

@@

The expansion of Medicaid reduced the cost of abortion and related care services, which may have increased the demand for abortion services. However, the expansion of Medicaid also increased access to contraception, which may have reduced the number of unwanted pregnancies and the demand for abortion services. Therefore, the impact of Medicaid expansion on abortion rates is ambiguous[@eliason2022]. To investigate the impact of Medicaid expansion on abortion rates, we used a difference-in-differences approach to compare changes in abortion rates in states that expanded Medicaid to those that did not.

@@

@@

## Impact of Telehealth Abortion on Abortion Rates

One in five abortion cases is through Telehealth. Telehealth abortion is a safe and effective way to access abortion services. Medication abortion accounted for 39% of all abortions in 2017. This number increased to 63% in 2020. Telehealth abortion has been shown to increase access to abortion services, reduce the abortion stigma.

@@

To investigate the impact of telehealth abortion on abortion rates, we used a difference-in-differences approach to compare changes in abortion rates in states that expanded telehealth abortion to those that did not.

@@

-   In April 2020, the F.D.A. removed[ the in-person requirement ](https://www.nytimes.com/2021/04/13/health/covid-abortion-pills-mailed.html "https://www.nytimes.com/2021/04/13/health/covid-abortion-pills-mailed.html")for the duration of the pandemic, allowing pills to be mailed.

-   On December 16^th^ 2021, the new F.D.A. decision made the suspension permanent.

-   On June 24, 2022, the US Supreme Court overturned Roe v. Wade.

@@

>>>>

---
title: "Idea Development"
date: 10/30/2024
date-modified: last-modified
date-format: full
categories: [Others]
website:
  comments: 
    utterances:
      repo: quarto-dev/quarto-web
#image: "cluster1.png"
tbl-colwidths: [5,30,125]
format: 
  html:
    page-layout: full
bibliography: references.bib
---

## Impact of Medicaid Expansion on Wait Time in the Hospitals

Timely treatment is crucial for healthcare outcome. The expansion of Medicaid lowered the barrier to receive healthcare service and could induce more patients to hospitals.

A compelling body of research have studies the impact of Medicaid expansion on the waiting time in the emergency room. [@wang2022] find that Medicaid expansion increases ED wait time by 10.4% (or 3.5 minutes) However, it is not clear how Medicaid expansion impact waiting time of non-urgent hospitals.

We conjecture that the expansion of Medicaid could attract patients withnon-urgent conditions to healthcare providers, yielding in an increase in waiting time for treatments.

## Impact of Covid-19 on Organ Donations

Organs from motor vehicle crashes is the largest organ transplant pool and accounts for **1 in 3** organ transplantation.

The stay-at-home rules significantly reduced the fatal traffic collisions, resulting in decrease in organ supply [@dembosky2020]. On the other hand, the pandemic also caused an decrease in public transportation usage and increase in personal vehicle usage [@ebrahimshaik2022]. The increase in personal vehicle usage could increase traffic collisions, which potentially leads to an increase in organ supply.

Our objective is examine the effects of Covid-19 on organ supplies.

## Impact of Medicaid Expansion on Length of Stay at Hospitals

@@

The expansion of Medicaid reduced the cost of abortion and related care services, which may have increased the demand for abortion services. However, the expansion of Medicaid also increased access to contraception, which may have reduced the number of unwanted pregnancies and the demand for abortion services. Therefore, the impact of Medicaid expansion on abortion rates is ambiguous[@eliason2022]. To investigate the impact of Medicaid expansion on abortion rates, we plan to use a difference-in-differences approach to compare changes in abortion rates in states that expanded Medicaid to those that did not.

[![](images/clipboard-1560225122.png){fig-align="center"}](https://www.pewresearch.org/short-reads/2024/03/25/what-the-data-says-about-abortion-in-the-us/sr_24-03-26_abortion_1-png/)

@@

**Expansion Timeline**

The following table shows the expansion timeline for Medicaid [@maclean2019a].

![](images/clipboard-665024476.png)

**Expansion Map**

@@

## Impact of Telehealth on Abortion Rates

1 in 5 abortion cases are through Telehealth in June 2024, which increased significantly compared to the 1 in 25 in April 2022 [@mulvihill2024]. Telehealth abortion is a safe and effective way to access abortion services. Medication abortion accounted for 39% of all abortions in 2017. This number increased to 63% in 2022 [@jones2024a]. Telehealth abortion has been shown to increase access to abortion services, reduce the abortion stigma.

@@

To investigate the impact of telehealth abortion on abortion rates, we plan to use a difference-in-differences approach to compare changes in abortion rates in states that expanded telehealth abortion to those that did not.

@@

-   In April 2020, the F.D.A. removed[ the in-person requirement ](https://www.nytimes.com/2021/04/13/health/covid-abortion-pills-mailed.html "https://www.nytimes.com/2021/04/13/health/covid-abortion-pills-mailed.html")for the duration of the pandemic, allowing pills to be mailed.
-   On December 16^th^ 2021, a new F.D.A. decision made the suspension permanent.
-   On June 24, 2022, the US Supreme Court overturned Roe v. Wade.

@@

**Issues**
